Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex 

### 2019-02-05 Manuscript.Rmd
Done with methods draft

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and $W$ is the adjusted correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected [@Fiedler1975]. Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks [@Fortunato2016].

In order to evaluate the validity of the resulting partitions, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the clustering that maximizes the network's modularity. Thus, unlike SP, modularity can fractionate a network into many functional groups. If the partitions from the bounded and unbounded community detection methods show high level of agreement, we can gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not.

@@

In order to evaluate the stability and topographical heterogeneity of the partitions within and across subjects, we chose the adjusted rand index (ARI) as the reference metric of agreement between two clusterings [@Hubert1985]. ARI estimates the rate at which every pair of nodes in partition A were jointly assigned to the same/different parcels on partition B, over all possible changes (e.g. clustered together on A, but separately on B). Because of this, the adjusted rand index determines the percentage of agreement between any two partitions while being agnostic to the labeling scheme. The resulting ratio is compared to a baseline given by the expectation of a random node assignment across partitions. This means that even though clustering deviations are heavily penalized, positive ARI values compare favorably against chance clustering (and can take negative values if the index falls below chance).

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject to validate the resulting functional patterns. Modularity and SP show overfitting and underfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018]. Their alignment would thus give us confidence in the validity of the resulting topologies. Next, we compared the functional organization derived from the whole time series among all subjects, and report the mean RI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there are mean differences in overall agreement within these regions by performing a paired permutation analysis (5000 iterations) on all pairwise RI comparisons across subjects. Given that the functional heterogeneity of brain areas is unevenly distributed across the cortex [@Mueller2013a], we perform the following analyses on PCC and mPFC separately.

To estimate the degree of stability of these partitions per subject we performed a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37), comparing each window's resulting organization against the clustering derived from the whole subject's data. Given the high degree of overlap across adjacent windows, we attribute a vertex's tendency to persistently shift allegiances to noise. In that regard, and consdiering the descriptive relevance of Fiedler eigenvector values in previous fMRI research [@Tian2018], we also examined the relationship between the magnitude of Fiedler vector values with the proportion of times a vertex was part of the DMN community across windows. To do this, we performed a mixed effects logistic regression model, with subject as a mixed effect and vertex-wise eigenmap values (derived from their full dataset) as fixed predictors on the proportion of times each vertex participated in the DMN community. Based on this relationship, we identified a threshold of Fiedler vector values for each subject, such that vertices above these thresholds were persistently associated with either DMN or non-DMN over 95% of the time. These thresholds allowed us to produce individual maps that retained the most stable vertices for further analyses, as well as discarding large sources of noise.

We then estimated the level of agreement between scans taken across days (i.e. pairs of daily left-right and right-left phase encoding sessions). If the functional organization estimated by SP is indeed subject-specific, we should see higher agreement for the same subject across days than either day compared to every other subject. We tested this idea by computing the ratio of the mean RI within and between subjects. Ratios close to one denoted good alignment across subjects, while ratios significantly higher than one suggested that partitions per subject were more stable than what you would expect against every other subject. We then extend this idea by computing the overall agreement across all sessions. Similar to the day-to-day analysis, we expect that sessions within subjects will show higher agreement than between subjects. Given what we know about these general regions, we expected PCC to be well aligned across participants, and stable within subjects. On the other hand, we hypothesized low agreement of mPFC partitions across participants, but with considerable stability across days and sessions within subject.

@@

BOLD signal recording through fMRI in mPFC in general, and vmPFC in particular, suffers from technical issues that promote signal dropout [@Logothetis]. Even though we made attempts to remove clear sources of noise through data preprocessing and stability analyses, it was still important to quantify the remaining potential influence of noise in our findings. We examined this through two complementary steps. First, we computed the temporal signal-to-noise ratio (tSNR) of all the vertices in the ROIs from HCP minimally preprocessed data (i.e. prior to the complementary denoising described in the "fMRI Data" section). We compared the mean tSNR for each region across subjects through a permutation analysis (all permutation analyses involved 5000 iterations). In addition, we compared the mean tSNR between mPFC and PCC through a permutation analysis for each separate subject, and report the percentage of subjects who had a significantly lower mean tSNR in mPFC after FDR correction for multiple testing [@Benjamini2001]. Next, we estimated which vertices within mPFC had a stronger relationship to the overall dynamics of DMN. To do this, we computed the mean time series of each subject's estimated DMN map, which was then predicted by the time series each mPFC vertex in a mass univariate analysis. The resulting $R^2$ values from each model were extracted to determine the total variance of the mean DMN time series explained by each individual vertex. If high noise in more ventral mPFC areas is affecting our results, we should see a gradient of $R^2$ values that decreases ventrally across subjects. We then spatially correlated (Pearson) the $R^2$ and tSNR values to determine whether low tSNR negatively impacted the individual contribution among vertices.

@@

A potential downfall of the present approach is that it can be computationally demanding, while more direct seed-based functional connectivity methods could provide sufficient approximations of an individual's DMN orgnization. We examined this issue by estimating mPFC functional maps from its vertex-wise functional coupling with the mean PCC activation. For each subject's concatenated day 1 and day 2 scans, we computed the mean time series of all vertices contained within the PCC region of a standard DMN arrangement [@Yeo2011], and correlated it (Pearons) with the time series of every mPFC vertex from our ROIs. The resulting day 1 and day 2 mPFC maps were then correlated within and between subjects in three ways: 1) between correlation-based maps; 2) between community-based maps; and 3) between correlation- and community-based maps. These spatial correlations are aimed to determine the day-to-day stability of each approach, as well as their level of agreement. These methods should produce somewhat similar arrangements, but the one displaying the highest within-subject agreement across days should be preferred (for example, see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

Finally, we examined whether each idiosyncrasies in functional topology corresponded to outstanding anatomical features. To do this, we correlated each subject's Fiedler vector values with their respective cortical thickness, cortical folding, and myelin density (quantified as the ratio of T1 to T2 magnitudes at every given vertex). These simple correlations are aimed to provide a coarse sense of this possible relationship, which can be studied in depth in the future.

>>>>

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlation values with every other vertex), and $W$ is the adjusted correlation matrix. This resulted in a matrix wherein each entry was the normalized value of the connection (from 0 to 1) between any two vertices relative to their combined connectivity strength, and with ones along the diagonal. This transormation ensures that every row adds up to zeo. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected [@Fiedler1975]. Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community to be the one containing the DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks [@Fortunato2016].

In order to evaluate the validity of the resulting partitions, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the clustering that maximizes the network's modularity. Thus, unlike SP, modularity can fractionate a network into many functional groups. We assumed that if the partitions from the bounded and unbounded community detection methods show high level of agreement, we could gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not.

@@

In order to evaluate the stability and topographical heterogeneity of the partitions within and across subjects, we chose the adjusted rand index (ARI) as the reference metric of agreement between two clusterings [@Hubert1985]. ARI estimates the rate at which every pair of nodes in partition A were jointly assigned to the same/different parcels on partition B, over all possible changes (e.g. clustered together on A, but separately on B). Therefore, the adjusted rand index determines the percentage of agreement between any two partitions while being agnostic to the labeling scheme. The resulting ratio is compared to a baseline given by the expectation of a random node assignment across partitions. This means that even though clustering deviations are heavily penalized, positive ARI values compare favorably against chance clustering (and can take negative values if the index falls below chance).

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject to validate the resulting functional patterns. Modularity and SP show overfitting and underfitting tendencies, respectively, in their community detection performance in a diverse number of network types [@Ghasemian2018]. Their alignment would thus give us confidence in the validity of the resulting topologies. Next, we compared the functional organization derived from the whole time series among all subjects, and report the mean ARI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there are mean differences in overall agreement within these regions by performing a paired permutation analysis (5000 iterations) on all pairwise RI comparisons across subjects. Given that the functional heterogeneity of brain areas is unevenly distributed across the cortex [@Mueller2013a], we performed all the subsequent analyses on PCC and mPFC separately.

To estimate the degree of stability of these partitions per subject we performed a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37), comparing each window's resulting partition against the clustering derived from the whole subject's data. Given the high degree of overlap across adjacent windows, we attribute a vertex's tendency to persistently shift allegiances to noise. In that regard, and consdiering the descriptive relevance of Fiedler eigenvector values in previous fMRI research [@Tian2018], we also examined the relationship between the magnitude of Fiedler vector values with the proportion of times a vertex was part of the DMN community across windows. To do this, we performed a mixed effects logistic regression model, with subject as a mixed effect and vertex-wise eigenmap values (derived from their full dataset) as fixed predictors on the proportion of times each vertex participated in the DMN community. Based on this relationship, we identified a threshold of Fiedler vector values for each subject, such that vertices above these thresholds were persistently associated with either DMN or non-DMN over 95% of the time. These thresholds yielded individual maps that retained the most stable vertices for further analyses, as well as discarding large sources of noise.

We then estimated the level of agreement between scans taken across days (i.e. pairs of daily left-right and right-left phase encoding sessions). If the functional organization estimated by SP is indeed subject-specific, we should see higher agreement for the same subject across days than either day compared to every other subject. We tested this idea by computing the ratio of the mean RI within and between subjects. Ratios close to one denoted good alignment across subjects, while ratios significantly higher than one suggested that partitions within a subject were more stable than what you would expect against every other subject. We then extend this idea by computing the overall agreement across all sessions. Similar to the day-to-day analysis, we expect that sessions within subjects would show higher agreement than between subjects. Given what we know about these general regions, we expected PCC to be fairly well aligned across participants, and stable within subjects. On the other hand, we hypothesized low agreement of mPFC partitions across participants, but with considerable stability across days and sessions within subject.

@@

BOLD signal recording through fMRI in mPFC in general, and vmPFC in particular, suffers from technical issues that promote signal dropout [@Logothetis]. Even though we made attempts to remove clear sources of noise through data preprocessing and stability analyses, it was still important to quantify the remaining potential influence of structured noise in our findings. We examined this through two complementary steps. First, we computed the temporal signal-to-noise ratio (tSNR) of all the vertices in the ROIs from the HCP minimally preprocessed data (i.e. prior to the complementary denoising described in the "fMRI Data" section). We compared the mean tSNR for each region across subjects through a permutation analysis (all permutation analyses involved 5000 iterations). In addition, we compared the mean tSNR between mPFC and PCC through a permutation analysis for each separate subject, and report the percentage of subjects who had a significantly lower mean tSNR in mPFC (post FDR correction for multiple testing [@Benjamini2001]).

Next, we examined whether a vertex's affiliation to the DMN community was driven by its tSNR. First, we computed the mean time series of each subject's DMN community. We then performed a mass univariate analysis to predict the average time series from each individual vertice's time series. The resulting $R^2$ values from each model were extracted to determine the total variance of the mean DMN time series explained by each individual vertex. If high noise in more ventral mPFC areas is affecting our results, we should see a gradient of $R^2$ values that decreases ventrally across subjects. We then spatially correlated (Pearson) each subject's $R^2$ and tSNR values to determine whether low tSNR negatively impacted the individual contribution among vertices, and report the mean correlation across subjects.

@@

A potential downfall of our approach lies in whether the considerable computational demands are indeed required to identify these individualized partitions. It is possible that careful seed-based functional connectivity could provide sufficiently stable approximations of an individual's DMN arrangement. Given the high variability and idiosyncrasy of mPFC (see Results), we examined if SP could capture more reliable maps than seed-based functional connectivity in this area. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average PCC activity for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by spatially correlating them across days and subjects in three ways: 1) between seed-based maps; 2) between community-based maps; and 3) between correlation- and community-based maps. These spatial correlations were aimed to determine the day-to-day stability of each approach, as well as the overall level of agreement between them. These methods should produce somewhat similar arrangements, but the one displaying the highest within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

The thresholded partitions in our work qualitatively resembled the topographical arrangement of networks A and B defined by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their work, Braga and Buckner selected individual vertices along the dorsolateral prefrontal cortex (DLPFC) that produced these two anticorrelated networks. Therefore, we hypothesized that if the communities we identified on the medial wall match their proposed networks, we should be able to find seed vertices in DLPFC that can reproduce them. The whole-brain seed-based functional connectivity maps for these subjects were qualitatively compared to those produced by spectral partitioning.

@@

Finally, we examined whether each idiosyncrasies in functional topology corresponded to outstanding anatomical features. To do this, we correlated each subject's thresholded Fiedler vector values with their respective cortical thickness, cortical folding, and myelin density (quantified as the ratio of T1 to T2 magnitudes at every given vertex). These simple correlations are aimed to provide a coarse sense of this possible relationship, which can be studied in depth in the future.
